---
title: "Project Draft"
date: 'Due: October 30, 11:59pm'
output:
  pdf_document: default
  html_document: default
authors: Alexis Bianco, Pierce Hollier, Hadeel Hamoud, Ryan Lee
---

```{r loading-pack-reading-data, message=FALSE}
library(tidyverse)
climate <- read_csv("data/WB_clim_migr_conf.csv")

#climate2 <- data.frame(t(climate[-1]))
#colnames(climate2) <- climate[, -1]

#glimpse(climate2)
```

### Introduction



### Methodology

In order to analyze our data, we will run multiple linear regressions on internally displaced people associated with disasters and other variables that have been associated with displacement. We will look at regions of countries in the world in order to make the data analysis easier for us. We would like to see which variables are correlated with internal displacement, and once we identify some of the better predictors of displacement, we would like to run other 

```{r}
displacement <- climate %>% 
  filter(`Series Name` == "Internally displaced persons, new displacement associated with disasters (number of cases)" | `Series Name` == "Rural population") %>% 
  filter(`Country Name` == "Vietnam") %>% 
  select(!(`Country Name` | `Country Code` | `Series Code`))




climate2 <- data.frame(t(displacement[-1]))
#colnames(climate2) <- displacement[, -1]

climate2 <- climate2 %>% 
  filter(X1 != ".." & X2 != "..")

disp_lm <- lm(X1 ~ X2, data = climate2)

tidy(disp_lm)



```


### Restults



### Discussion


